---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

  library(magrittr)

```

# `envClean`: an R package to help clean large unstructured data set(s)

<!-- badges: start -->
<!-- badges: end -->

The goal of `envClean` is to help clean large amounts of unstructured, biological data (for further analysis elsewhere).

Not all functions will be relevant to all projects.

If a typical species list from a typical observer is required, then `make_effort_mod()` may be useful to filter out excessively rich or depauperate lists.

If many data sources are included in the incoming data, taxonomic alignment via `make_taxonomy()` is likely to be required. If those data sources are likely to contain duplicates, using taxonomic, geographic and temporal bins may be the easiest way to ensure duplicates are removed.

Some functions could be considered 'experimental'. `add_cover()` uses principal components analysis on environmental variables to generate a best guess for percentage cover where some records are missing that attribute.

## Installation

`envClean` is not on [CRAN](https://CRAN.R-project.org).

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Acanthiza/envClean")
```

Load `envClean`

```{r}
library("envClean")
```

## Other packages/resources

These are unrelated to `envClean` and are possibly much better documented:

* The book [Cleaning Biodiversity Data in R](https://cleaning-data-r.ala.org.au/)
* R packages:
    + [CoordinateCleaner](https://cran.r-project.org/web/packages/CoordinateCleaner/vignettes/Cleaning_GBIF_data_with_CoordinateCleaner.html)
    + [bdc](https://cran.r-project.org/web/packages/bdc/index.html)

## What is in `envClean`

The following functions and data sets are provided in `envClean`. See https://acanthiza.github.io/envClean/ for more examples.

```{r, echo = FALSE, warning = FALSE}

  get_name_title <- function(text) {
    
    tibble::enframe(text
                    , name = NULL
                    , value = "description"
                    ) %>%
      dplyr::filter(grepl("title", description)) %>%
      dplyr::mutate(description = gsub("\\\\|title", "", description)
                    , description = gsub("^\\{|\\}$", "", description)
                    )
    
  }

  manuals <- fs::dir_info("man") %>%
    dplyr::select(path) %>%
    dplyr::filter(grepl("Rd$", path)
                  , !grepl("pipe", path)
                  ) %>%
    dplyr::mutate(object = gsub("man\\/|\\.Rd","",path)
                  , class = purrr::map_chr(object, ~envFunc::vec_to_sentence(class(get(.))))
                  , text = purrr::map(path, readLines)
                  , text = purrr::map(text, get_name_title)
                  , object = paste0("envClean::"
                                    , object
                                    , ""
                                    )
                  , object = dplyr::if_else(class == "function"
                                            , paste0(object
                                                     , "()"
                                                     )
                                            , object
                                            )
                  , object = paste0("`", object, "`")
                  ) %>%
    tidyr::unnest(cols = c(text))

  knitr::kable(manuals[,2:4])

```




