---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

  library(magrittr)

```

# `envClean`: an R package to help clean large unstructured data set(s)

<!-- badges: start -->
<!-- badges: end -->

The goal of `envClean` is to help clean large, unstructured, biological (or `env`ironmental) data sets.

## Installation

`envClean` is not on [CRAN](https://CRAN.R-project.org).

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Acanthiza/envClean")
```

Load `envClean`

```{r}
library("envClean")
```

## Filtering an 'area of interest'

An area of interest, or geographic range, sets the spatial boundary for the cleaning. Adding geographic (or spatial) bins based on a raster that spans the area of interest is another way to achieve the same end.

This example uses the `flor_all` data frame and the simple feature `aoi`. Converting `flor_all` to `sf` allows plotting them together.

Load `flor_all`

```{r}
flor_all <- tibble::as_tibble(flor_all)
```

Convert `flor_all` to `sf` and plot together with `aoi`.

```{r aoi, fig.cap = "Records from `flor_all` plotted over the area of interest `aoi`."}

  flor_all_sf <- flor_all %>%
    sf::st_as_sf(coords = c("long", "lat")
                 , crs = 4326
                 )

  tmap::tm_shape(aoi
           , bbox = sf::st_bbox(flor_all_sf)
           ) +
    tmap::tm_polygons() +
  tmap::tm_shape(flor_all_sf) +
    tmap::tm_dots()

```

Filtering `flor_all` to `aoi` is done with `filter_geo_range`.

```{r flor_all_aoi}

  flor_aoi <- filter_geo_range(flor_all
                         , use_aoi = aoi
                         ) %>%
    envFunc::add_time_stamp()

  flor_aoi

```

Check that spatial filter worked.

```{r flor_aoi}

  flor_aoi_sf <- flor_aoi %>%
    sf::st_as_sf(coords = c("long", "lat")
                 , crs = 4326
                 )

  tmap::tm_shape(aoi
           , bbox = sf::st_bbox(flor_all_sf)
           ) +
    tmap::tm_polygons() +
  tmap::tm_shape(flor_aoi_sf) +
    tmap::tm_dots()

```

## What else is in `envClean`

The following functions and data sets are provided in `envClean`. See https://acanthiza.github.io/envClean/ for more examples.

```{r, echo = FALSE, warning = FALSE}

  get_name_title <- function(text) {
    
    tibble::enframe(text
                    , name = NULL
                    , value = "description"
                    ) %>%
      dplyr::filter(grepl("title", description)) %>%
      dplyr::mutate(description = gsub("\\\\|title", "", description)
                    , description = gsub("^\\{|\\}$", "", description)
                    )
    
  }

  manuals <- fs::dir_info("man") %>%
    dplyr::select(path) %>%
    dplyr::filter(grepl("Rd$", path)
                  , !grepl("pipe", path)
                  ) %>%
    dplyr::mutate(object = gsub("man\\/|\\.Rd","",path)
                  , class = purrr::map_chr(object, ~envFunc::vec_to_sentence(class(get(.))))
                  , text = purrr::map(path, readLines)
                  , text = purrr::map(text, get_name_title)
                  , object = paste0("envClean::"
                                    , object
                                    , ""
                                    )
                  , object = dplyr::if_else(class == "function"
                                            , paste0(object
                                                     , "()"
                                                     )
                                            , object
                                            )
                  , object = paste0("`", object, "`")
                  ) %>%
    tidyr::unnest(cols = c(text))

  knitr::kable(manuals[,2:4])

```




