### Percentage of sites

```{r kept}

  kept <- flor_prop %>%
    dplyr::mutate(cells = n_distinct(across(all_of(visit_cols)))) %>%
    dplyr::filter(taxa %in% keep_taxa) %>%
    dplyr::count(taxa,cells) %>%
    dplyr::mutate(per = round(100*n/cells,2))

  min_per_kept <- min(kept$per)


```

In order for a taxa to be retained, it needed to be present in more than `r min_per_kept`% of contexts (or `r round(nrow(flor_prop %>% dplyr::distinct(across(any_of(visit_cols)))) * min_per_kept/100,0)` contexts). This percentage was determined by looking for the minimum value required to include several key structural species with few contexts (e.g. `r vec_to_sentence(paste0("_",kept$taxa,"_"))`. See Table \@ref(tab:sitePerFilterTaxa)).

Filtering was done with the `filter_prop` function in the `envClean` package `r cite_package("envClean")`.

`r data_filter$filter_text[i]`

<br>

```{r sitePerFilterTaxa}

   kept %>%
    dplyr::select(Taxa = taxa,  contexts = n, `_Percentage_ of sites` = per) %>%
    kable(caption = "Number and percentage of contexts at which several key taxa were recorded")

```

<br>
