### Taxonomy

```{r remove}

  regexRemove <- c("quotation marks", "question marks", "trailing 'spp.' or 'sp.'", "leading space", "trailing space")

```

Taxa were queried against the global biodiversity information facility (GBIF) [taxonomy backbone](https://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c) using the `make_taxa_taxonomy` and `filter_taxa` functions in the `envClean` package `r cite_package("envClean")`. These functions rely heavily on the `name_backbone` function in the `rgbif` package `r cite_package("rgbif")`.

Prior to querying against GBIF the following cleaning steps were undertaken:

* any record of a taxa that contained any of the following text in its name (ignoring case) was removed:
    + `r paste0(sort(unique(species_filt)), collapse = "\n    + ")`
* any of the following regular expressions in a taxa name were removed from the name (via `gsub`):
    + `r paste0("'", na.omit(gsub("^c|\\(|\\)$", NA, formals(get_gbif_tax)$remove_strings)), "'", collapse = "\n    + ")`
* any text including and after ' x ' or ' X ', in a taxa name, was removed from the name (thus any hybrids were most likely assigned to the first taxa recorded within the hybrid)

The taxa used in analyses were always binomial (taxa recorded at taxonomic level below species were grouped at species level (e.g. _Acanthiza iredalei rosinae_ was changed to _Acanthiza iredalei_). Any record that could not be matched to a name at `r rank_to_target` level, or better, was removed. Table \@ref(tab:taxHier) shows each taxonomic rank and the count of taxa matched at that rank.

At the same time as the analysis-wide taxonomy was applied (by `filter_taxa`), any duplication of cover and lifeform for a taxa in a context was resolved, leaving only a single record of each taxa within a context.

`r data_filter$filter_text[i]`

The [appendix] Table \@ref(tab:taxonomyApp) maps names retrieved from data sources to those used in analyses.

<br>

```{r taxHier}

  kable(lutaxa %>%
          dplyr::inner_join(filt_summary$obj[[1]] %>%
                              dplyr::count(original_name, name = "records")
                            ) %>%
          dplyr::group_by(rank) %>%
          dplyr::summarise(taxa = n_distinct(original_name)
                           , Records = sum(records)
                           ) %>%
          dplyr::ungroup() %>%
          dplyr::arrange(desc(rank))
        , caption = "Count of taxa (and records) identified to various taxonomic rank"
        )

```

<br>
