### Proportion cover

```{r flor_bdbsa}

  flor_bdbsa <- flor_aoi %>%
    dplyr::filter(data_name == "BDBSA")

```

During import from the original data sources (proportion) cover was derived in different ways depending on the available information:

* for BDBSA, the `COVCODE` or `COVER` fields were used
  + `COVCODE` stores a modified Braun-Blanquet [@RN4265] cover value (see table \@ref(tab:bbCover))
  + `COVER`, which was rarely used (`r round(100*sum(flor_aoi$cover > 0 & flor_aoi$cover <= 100,na.rm=TRUE)/nrow(flor_aoi),1)`% of records), stores a percentage cover value
* for TERN, the proportion of sampling points that were a taxa
* for ALIS, each crown separation ratio was mapped to a Braun-Blanquet cover value then treated as per BDBSA COVCODE (see table \@ref(tab:csrCover))
* for BCM, no cover estimates were available at taxa level (cover estimates were recorded for structural layers only)
* for PTP, 'sites' were generated within polygons developed by that project by taking the point furthest away from a polygon boundary. The taxa list, including cover estimates for the polygon was then applied to that 'cell'
* for NVB, no estimates of cover were available

Cover estimates were then assigned to all records:
  + if a field-based cover estimate was available, that was used
  + if there was no field-based cover estimate, the median cover value for that taxa within [pca groups] was used
  + if there was still no cover estimate available the taxa was assigned a small cover value (`r min(lucover %>% dplyr::pull(cov_type))`)
  
This was done using the `make_cover` function in the `envClean` package `r cite_package("envClean")`.
      
If all taxa within a context required their cover estimate to be assigned as a small value, that context was removed.

`r data_filter$filter_text[i]`
      
<br>

```{r bbCover}

  lucover %>%
    dplyr::select(Code = cover_code
                  , Description = cover_description
                  , `Proportion cover value used` = !!ensym(cov_type)
                  ) %>%
    kable(caption = "Translation of recorded Braun-Blanquet values to proportion cover values used")

```

<br>

```{r csrCover}

  luCSR %>% 
    dplyr::select(`Crown separation ratio recorded` = CSR
                  , Description = CSR_Description
                  , `Braun-Blanquet value used` = COVCODE
                  ) %>%
    kable(caption = "Translation of recorded crown separation ratio to modified Braun-Blanquet values")


```

<br>
